---
title: "Align different batch peak tables"
author:
- name: Xiaotao Shen (https://www.shenxt.info/)
date: "Created on 2020-04-01 and updated on `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: no
  pdf_document:
    toc: no
vignette: >
  %\VignetteIndexEntry{align_batch}
  %\VignettePackage{masscleaner}
  % \VignetteEngine{knitr::rmarkdown}
  % \usepackage[utf8]{inputenc}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = TRUE,
  out.width = "100%"
)
```

Some times if you have more than two batch peak tables, and they are processed at different times, so you need to align them together before other analysis. In `masscleaner`, we use the `align_batch()` function to do that.

# **Data preparation**

See the [massdataset](https://tidymass.github.io/massdataset/) package, and create you metabolomics dataset into 2 mass_dataset objects.

Here we use the demo data from `demodata` package, so please install it first.

```{r,eval=FALSE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
if(!require(devtools)){
install.packages("devtools")
}
devtools::install_github("tidymass/demodata")
```


```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
library(masscleaner)
library(demodata)
library(tidyverse)
```


# **Run `align_batch()` function**

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
data(object1, package = "demodata")
data(object2, package = "demodata")

object1
object2

x = object1
y = object2

match_result =
  align_batch(x = object1, y = object2, return_index = TRUE)

head(match_result)

new_object =
  align_batch(x = object1, y = object2, return_index = FALSE)

new_object
```


# **Session information**

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
sessionInfo()
```






