---
title: "Data normalization and integration"
author:
- name: Xiaotao Shen PhD (https://www.shenxt.info/)
date: "Created on 2022-01-18 and updated on `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: no
  pdf_document:
    toc: no
vignette: >
  %\VignetteIndexEntry{data_normalization_integration}
  %\VignettePackage{masscleaner}
  % \VignetteEngine{knitr::rmarkdown}
  % \usepackage[utf8]{inputenc}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = TRUE,
  out.width = "100%"
)
```

We can use `masscleaner` for data normalization and data integration.

First, we need to prepare samples for `masscleaner`.

# **Data preparation**

Load data from [previous step](https://tidymass.github.io/masscleaner/articles/missing_value_imputation.html).

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUE, cache=TRUE}
library(masscleaner)
library(tidyverse)
load("peak_tables/POS/object")
object
```

# **Data quality assessment**

We can use the [`massqc` package](https://tidymass.github.io/massqc/) to assess the data quality.

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUE, cache=TRUE}
library(massqc)
object <-
  object %>% 
  activate_mass_dataset(what = "sample_info") %>% 
  dplyr::mutate(batch = as.character(batch))

object %>% 
  `+`(1) %>% 
  log(2) %>% 
massqc::massqc_pca(color_by = "batch", line = FALSE)
```

We can see the clear batch effect.

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUE, cache=TRUE}
object %>% 
  activate_mass_dataset(what = "sample_info") %>% 
  dplyr::filter(class == "QC") %>% 
  `+`(1) %>% 
  log(2) %>% 
  massqc::massqc_sample_boxplot(color_by = "group", 
                                order_by = "injection.order") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUE, cache=TRUE}
object %>% 
  activate_mass_dataset(what = "sample_info") %>% 
  dplyr::filter(class == "QC") %>% 
  `+`(1) %>% 
  log(2) %>% 
  massqc::massqc_rsd_plot()
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUE, cache=TRUE}
object %>% 
  activate_mass_dataset(what = "sample_info") %>% 
  dplyr::filter(class == "QC") %>% 
  `+`(1) %>% 
  log(2) %>% 
  massqc::massqc_sample_correlation(method = "square", tl.cex = 5) +
  theme(axis.text = element_text(size = 5))
```

# **Data normalization**

## Total/median/mean

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUE, cache=TRUE}
object1 <- 
  normalize_data(object, method = "total")

##PCA
object1 %>% 
  `+`(1) %>% 
  log(2) %>% 
massqc::massqc_pca(color_by = "batch", line = FALSE)
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUE, cache=TRUE}
object2 <- 
  normalize_data(object, method = "median")

##PCA
object2 %>% 
  `+`(1) %>% 
  log(2) %>% 
massqc::massqc_pca(color_by = "batch", line = FALSE)
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUE, cache=TRUE}
object3 <- 
  normalize_data(object, method = "mean")

##PCA
object3 %>% 
  `+`(1) %>% 
  log(2) %>% 
massqc::massqc_pca(color_by = "batch", line = FALSE)
```

## Loess/SVR

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUE, cache=TRUE}
object4 <- 
  normalize_data(object, method = "loess")

##PCA
object4 %>% 
  `+`(1) %>% 
  log(2) %>% 
massqc::massqc_pca(color_by = "batch", line = FALSE)
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUE, cache=TRUE}
object5 <- 
  normalize_data(object, method = "svr")

##PCA
object5 %>% 
  `+`(1) %>% 
  log(2) %>% 
massqc::massqc_pca(color_by = "batch", line = FALSE)
```

We select the median method.


```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUE, cache=TRUE}
object_normalization <- object2
```

# **Data integration**

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUE, cache=TRUE}
object_integration <- 
  integrate_data(object_normalization, method = "qc_median")

##PCA
object_integration %>% 
  `+`(1) %>% 
  log(2) %>% 
massqc::massqc_pca(color_by = "batch", line = FALSE)
```

Then we can draw the compare plot for RSD:

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUE, cache=TRUE}
qc_id = 
  object %>% 
  activate_mass_dataset(what = "sample_info") %>% 
  filter(class == "QC") %>% 
  pull(sample_id)

rsd_before = 
object %>% 
  mutate_rsd(according_to_samples = qc_id) %>% 
  activate_mass_dataset(what = "variable_info") %>% 
  pull(rsd)

rsd_after = 
object_integration %>% 
  mutate_rsd(according_to_samples = qc_id) %>% 
  activate_mass_dataset(what = "variable_info") %>% 
  pull(rsd)


data.frame(rsd_before, rsd_after) %>% 
  dplyr::mutate(class = dplyr::case_when(rsd_after < rsd_before ~ "Decrease",
                                         rsd_after > rsd_before ~ "Increase",
                                         rsd_after == rsd_before ~ "Equal")) %>% 
  ggplot(aes(rsd_before, rsd_after, colour = class)) +
  ggsci::scale_color_jama() +
  geom_abline(slope = 1, intercept = 0) +
  geom_point() +
  labs(x = "RSD after normalization", y = "RSD before normalization") +
  theme_bw()
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUE, cache=TRUE}
intensity_plot(object = object, variable_index = 1, 
               color_by = "group", 
               order_by = "injection.order", interactive = FALSE) 
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUE, cache=TRUE}
intensity_plot(object = object_integration, variable_index = 1, 
               color_by = "group", 
               order_by = "injection.order", interactive = FALSE) +
  ggsci::scale_fill_aaas()
```


# **Session information**

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
sessionInfo()
```
