---
title: "Missing value imputation"
author:
- name: Xiaotao Shen (https://www.shenxt.info/)
date: "Created on 2020-04-01 and updated on `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: no
  pdf_document:
    toc: no
vignette: >
  %\VignetteIndexEntry{missing_value_imputation}
  %\VignettePackage{masscleaner}
  % \VignetteEngine{knitr::rmarkdown}
  % \usepackage[utf8]{inputenc}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = TRUE,
  out.width = "100%"
)
```

# **Introduction**

We can use `masscleaner` for missing value (MV) imputation.

First, we need to prepare samples for `masscleaner`.

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUEE, cache=TRUE}
library(masscleaner)
library(massdataset)
library(tidyverse)
```

# **Data preparation**

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUEE, cache=TRUE}
data("expression_data")
data("sample_info")
data("variable_info")
object =
  create_mass_dataset(
    expression_data = expression_data,
    sample_info = sample_info,
    variable_info = variable_info
  )
object

get_mv_number(object)
massdataset::get_mv_number(object, by = "sample")
head(massdataset::get_mv_number(object, by = "variable"))

massdataset::get_mv_number(object, by = "sample", show_by = "percentage")
head(massdataset::get_mv_number(object, by = "variable"), show_by = "percentage")
```

# **Filter noisy features**

Remove variables who have mv in more than 20% QC samples.

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUEE, cache=TRUE}
qc_id =
  object %>%
  activate_mass_dataset(what = "sample_info") %>%
  filter(class == "QC") %>%
  pull(sample_id)

subject_id =
  object %>%
  activate_mass_dataset(what = "sample_info") %>%
  filter(class == "Subject") %>%
  pull(sample_id)

object =
  object %>%
  mutate_variable_na_freq(according_to_samples = qc_id) %>%
  mutate_variable_na_freq(according_to_samples = subject_id) %>%
  activate_mass_dataset(what = "variable_info") %>%
  filter(na_freq < 0.2 & na_freq.1 < 0.5)
object
```

# **Filter outlier samples**

Remove samples with MV > 50% except Blank samples.

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUEE, cache=TRUE}
object =
  filter_samples(
    object = object,
    flist = function(x) {
      sum(is.na(x)) / nrow(object) < 0.5
    },
    apply_to = c(qc_id, subject_id),
    prune = TRUE
  )
```

# **Missing value processing**

For Blank samples and subject samples, we use different method to impute missing
values.

For Blank samples, just use the zero.

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUEE, cache=TRUE}
blank_id =
  object %>%
  activate_mass_dataset(what = "sample_info") %>%
  filter(class == "Blank") %>%
  pull(sample_id)

object1 =
  impute_mv(object = object,
            sample_id = blank_id,
            method = "zero")

object1 %>%
  activate_mass_dataset(what = "expression_data") %>%
  select(dplyr::contains("Blank")) %>%
  extract_expression_data() %>% 
  head()
```

For non-Blank samples, just use the knn method

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUEE, cache=TRUE}
non_blank_id =
  object %>%
  activate_mass_dataset(what = "sample_info") %>%
  filter(class != "Blank") %>%
  pull(sample_id)

non_blank_id

object2 =
  impute_mv(object = object1,
            sample_id = non_blank_id,
            method = "knn")

object2 %>%
  activate_mass_dataset(what = "sample_info") %>%
  filter(class != "Blank") %>%
  extract_expression_data() %>% 
  head()

object2 %>%
  activate_mass_dataset(what = "sample_info") %>%
  filter(class == "Blank") %>%
  extract_expression_data() %>% 
  head()
```


# **Session information**

```{r,eval=TRUE,warning=FALSE, R.options="", message=TRUEE, cache=TRUE}
sessionInfo()
```
